### 2021-03-03 article/paper_eng.tex
update

% \listfiles
\documentclass[a4paper]{article}
\usepackage[T2A]{fontenc}
\usepackage[english,russian]{babel}
\usepackage{sectsty}
\sectionfont{\centering\normalfont}
\subsectionfont{\centering\normalfont}
\subsubsectionfont{\normalfont}
\usepackage{changepage}
\usepackage{filecontents}
% \usepackage{mathptmx}
\usepackage{float} 
\usepackage{subcaption}
\usepackage[utf8]{inputenc}
\usepackage[14pt]{extsizes}
\usepackage{physics} 
\usepackage{graphicx}
\usepackage{enumerate}
\graphicspath{{../misc/}}
\usepackage{setspace,amsmath}
% \usepackage[left=30mm, right=15mm, top=20mm, bottom=20mm]{geometry} 
\usepackage[nottoc]{tocbibind}
\usepackage[style=gost-numeric,sorting=none,backend=biber,defernumbers=false]{biblatex}
\DeclareBibliographyAlias{misc}{article}
\addbibresource{lit.bib} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \everymath{\displaystyle}
\DeclareMathSizes{12}{20}{14}{10}

@@

The introduction of inhibitory connections (having negative weights) allows to achieve a better distinction of features learned by the neurons. The most common way to introduce competition between SNN neurons is to use a group of winner-take-all methods, where the first neuron in the layer that generates an impulse either zeroes the membrane variables of other neurons or prohibits other neurons from generating spikes for a given time. The use of inhibitory connections in the form of negative synaptic weights has not been studied either, and the influence of the possibility of updating them according to local rules, along with stimulating projections, has been hardly studied too. In this work, it is shown that learn-ing competition relationships in the architecture of a locally connected SNN potentially allows to achieve better results in the accuracy of pattern recognition (on MNIST digits) due to the organization of competition and cooperation between groups of neurons.

@@

The neural network architecture of LCSNN is inspired by the structure of the visual cortex of the brain. The $Y$ network layer consists of $n$ channels, each of which is locally connected to the $X$ layer of neurons. We connect neurons with common receptive fields by inhibitory connections.

\begin{figure} \label{LCSNN}
\includegraphics[,
 width=\textwidth,keepaspectratio=true]{LCSNN_ru.pdf} 
    \caption{LCSNN architecture}
\end{figure}

@@

Thus, in the process of learning, the weights of connections where pre-spikes are systematicly recieved right before the emmision of post-spikes are increased, and decreased if vice versa. After such training, the neurons begins to actively react to pre-spikes received from neurons with larger connection weights. If these pre-spikes are recieved in a short enough time period, they might cause a new spike. The opposite rule (with $A_{+}$ and $A_{-}$ having the opposite signs) is called the anti-STDP \cite{anti-STDP} rule. It is used to train inhibitory connenctions in section \ref{section:competition-training} of this paper.

@@

\begin{figure}
\centering
\begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{weights_XY_untrained_ru.pdf}
    \caption{Before training}
\end{subfigure}
\begin{subfigure}{0.45\textwidth}  \label{weights_XY}
    \includegraphics[width=\textwidth,keepaspectratio=true]{weights_XY_ru.pdf}
    \caption{After 1000 training iterations}
\end{subfigure}
\caption{Visualization of $XY$ connection weights of a 900 $Y$ neurons (100 neurons for each of the 9 patches) network --- a 360 $ \times $ 360 matrix. The weights of one neuron correspond to a 12 $\times$ 12 square. The weights of the $Y$ neurons are grouped by patch - the lower left 120 $  \times $ 120 square  (10 $ \times $ 10 neurons) corresponds to the lower left patch, the center square corresponds to the center patch, and so on. It can be seen that after training, the neurons learn features where digits elements can be observed.}
\end{figure}

@@

For all networks, this calibration process was used for 10,000 examples from a calibration sample especially selected from the test part of the MNIST dataset, as noted above. Note that calibration is nota  part of the network training, since it is only included in the algorithm for interpreting the behavior of the SNS. The votes are used as a measure of confidence of the neuron in each of the digits classes.

\begin{figure}
\centering
\begin{subfigure}{\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{votes_ru.pdf}
    \caption{$Y$ neurons votes.}
\end{subfigure}
\begin{subfigure}{\textwidth} 
    \includegraphics[width=\textwidth,keepaspectratio=true]{votes_distr_4_ru.pdf}
    \caption{Votes distrivution for the 4 most relevant classes (for each neuron). It can be seen that the vote value (average number of spikes per class) drops significantly with relevance decreasing.}
\end{subfigure}
\caption{Visualization of the votes of the $ Y $ network layer with 225 $ Y $ neurons. High values correspond to a large specialization of the neuron in the digits class. It can be seen that almost all neurons have a nonzero vote value, therefore, almost all neurons in the network are active.}
\end{figure}

@@

Accuracy learning curves were built for various interpretation algorithms. The accuracy was measured every 250 training iterations. To calibrate the interpretation algorithm at each step a calibration sample of 5000 was used, while to measure the accuracy, a test sample of 1000 was used.

@@

\begin{figure}
\centering
\begin{subfigure}{0.48\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{LCSNN_learning_rate_ru.pdf}
 \caption{LCSNN learning curve.}
 \label{LCSNN_learning_curve}
\end{subfigure} 
\begin{subfigure}{0.48\textwidth} 
    \includegraphics[width=\textwidth,keepaspectratio=true]{CSNN_learning_rate_ru.pdf}
 \caption{CSNN learning curve.}
 \label{CSNN_learning_curve}
\end{subfigure} 
\end{figure}
\begin{figure}\ContinuedFloat
\begin{subfigure}{0.48\textwidth} 
    \includegraphics[width=\textwidth,keepaspectratio=true]{FCSNN_learning_rate_ru.pdf}
 \caption{FCSNN learning curve.}
 \label{FCSNN_learning_curve}
\end{subfigure}
\caption{Comparison of learning curves of different architectures of SNNs with competition: (\ref{LCSNN_learning_curve}) LCSNN, (\ref{CSNN_learning_curve}) CSNN, (\ref{FCSNN_learning_curve}) FCSNN. The results are evaluated with different algorithms for the interpretation of activity. The \textbf{preselections by spikes} algorithm was only checked for LCSNN, because it achieves practically same results as \textbf{patch voting}. Standard deviations are plotted as errors. The networks have 100 channels (for a FCSNN the channel consists of a single neuron), the convolution kernel for LCSNN is 12, and for CSNN it is 8.}
\end{figure}

\subsection{Comparison of convolution and local connections effectiveness}
Были проведены эксперименты по измерению точности сетей с различными архитектурами. Для сверточных и полносвязных сетей вводились аналогичные LCSNN связи конкуренции. Из-за высоких вычислительных нагрузок не ставилось задачи по нахождению параметров, обеспечивающих максимальную точность для каждой архитектуры. Эти параметры были подобраны приблизительно, однако, судя по результатам отдельных отладочных экспериментов, величина расхождения с оптимальными значениями не превышает 1-2\%. Заметим, что при использовании линейного классификатора в качестве алгоритма интерпретации удалось достигнуть 95\% точности для локально соединенной сети из 1000 каналов с размером ядра 12.

Experiments were carried out to measure the accuracy of networks with different architectures. For convolutional and fully connected networks,  inhibitory connections similar to LCSNN were introduced. Note that using a linear classifier as an interpretation algorithm, it was possible to achieve 95\% accuracy for a locally connected network of 1000 channels with kernel size of 12.

@@

Следует отметить, что SNN могут достигать значительно больших точностей распознавания на MNIST. Для этого можно использовать: (i) cети с существенно большим числом весовых параметров (например, с увеличенным числом каналов); (ii) более глубокие сети (с большим количеством слоев); (iii) более эффективные на сегодня алгоритмы обучения (например, обучение с учителем или с использованием контрастивной функции потерь \cite{contrastive_loss}). В то же время, достижение максимально возможной точности не являлось целью настоящего исследования, адресованного, в первую очередь, сравнению различных архитектур SNN.

@@

\begin{figure}
\centering
\begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_distribution_worst_ru.pdf}
    \caption{Weakest competition}
\end{subfigure}
\begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_distribution_medium_bad_ru.pdf}
    \caption{Weak competition}
\end{subfigure}
\\
\begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_distribution_medium_good_ru.pdf}
    \caption{Medium competition} 
\end{subfigure}
\begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_distribution_best_ru.pdf}
    \caption{Strong competition}
    \label{fig:best_competition}
\end{subfigure}
\caption{Different competition weights distributions}
\label{fig:competition_distributions}
\end{figure}

@@

\begin{figure}
\centering
\begin{subfigure}{0.45\textwidth} 
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_on_XY_worst_ru.pdf}
    \caption{Weakest competition}
    \label{fig:worst_competition_distribution}
\end{subfigure}
\begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_on_XY_medium_bad_ru.pdf}
    \caption{Weak competition}
    \label{fig:medium_bad_competition_distribution}
\end{subfigure}
\\
\begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_on_XY_medium_good_ru.pdf}
    \caption{Medium competition}
    \label{fig:medium_good_competition_distribution}
\end{subfigure}
\begin{subfigure}{0.45\textwidth} 
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_on_XY_best_ru.pdf}
    \caption{Strong compettion}
    \label{fig:best_competition_XY}
\end{subfigure}
\caption{Visualization of competition weights on top of $ XY $ weights for the networks from Fig. (\ref{fig:competition_distributions}). Competition weights are shown only for one center neuron in each receptive field to avoid cluttering the visualization. The saturated blue color corresponds to the competition weights of large absolute value (the arithmetic mean between the weights $ W_{ij} $ and $ W_{ji} $ is used). It can be seen that similar features compete more strongly with each other than different}
\end{figure}

@@

\begin{figure}
\begin{center}
 \includegraphics[,
 width=\textwidth,keepaspectratio=true]{hyperparams_ru.png}

@@

\begin{figure}
\centering 
\begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_distribution_clamp_low_ru.pdf}
    \caption{Inhibitory connections weights distribution, lower clamp} 
\end{subfigure}
\begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_distribution_clamp_high_ru.pdf}
    \caption{Inhibitory connections weights distribution, higher clamp}
\end{subfigure}
\begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{weights_XY_clamp_low_ru.pdf}
    \caption{Forward connections weights, lower clamp}
\end{subfigure} 
\begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{weights_XY_clamp_high_ru.pdf}
    \caption{Forward connections weights, higher clamp}
\end{subfigure} 
\caption{Impact of limiting the values of inhibitory weights on accuracy. The weights were limited to $ -350 $ on the illustration below and $ -100 $ on the illustration above. The accuracy of a network with similar parameters, but without competition limitation is 0.823, its weights $ XY $ are shown in Fig. \ref {fig:best_competition_XY}, and the distribution of its inhibitory weights is shown in Fig. \ref{fig:best_competition}. It can be seen that in both cases more blurred forward propagation weights are trained.
}
\label{fig:compe_clamp}
\end{figure}

>>>>

% \listfiles
\documentclass[a4paper]{article}
\usepackage[T2A]{fontenc}
% \usepackage[english, russian]{babel}
\usepackage{sectsty}
\sectionfont{\centering\normalfont}
\subsectionfont{\centering\normalfont}
\subsubsectionfont{\normalfont}
\usepackage{changepage}
\usepackage{filecontents}
% \usepackage{mathptmx}
\usepackage{float} 
\usepackage{subcaption}
\usepackage[utf8]{inputenc}
\usepackage[14pt]{extsizes}
\usepackage{physics} 
\usepackage{graphicx}
\usepackage{enumerate}
\graphicspath{{../misc/}}
\usepackage{setspace,amsmath}
% \usepackage[left=30mm, right=15mm, top=20mm, bottom=20mm]{geometry} 
\usepackage[nottoc]{tocbibind}
\usepackage[style=gost-numeric,sorting=none,backend=biber,defernumbers=false]{biblatex}
\DeclareBibliographyAlias{misc}{article}
\addbibresource{lit.bib} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \everymath{\displaystyle}
\DeclareMathSizes{12}{20}{14}{10}

@@

The introduction of inhibitory connections (having negative weights) allows to achieve a better distinction of features learned by the neurons. In this work, it is shown that learning competition relationships in the architecture of a locally connected SNN potentially allows to achieve better results in the accuracy of pattern recognition (on MNIST digits) due to the organization of competition and cooperation between groups of neurons.

@@

The neural network architecture of LCSNN is inspired by the structure of the visual cortex of the brain. The $Y$ network layer consists of $n$ channels, each of which is locally connected to the $X$ layer of neurons, and neurons with common receptive fields are connected by inhibitory connections (Fig. \ref{fig:LC_SNN_architecture}).

\begin{figure} \label{LCSNN}
\includegraphics[,
 width=\textwidth,keepaspectratio=true]{LCSNN.pdf} 
    \caption{LCSNN architecture}
    \label{fig:LC_SNN_architecture}
\end{figure}

@@

Thus, in the process of learning, the weights of connections where pre-spikes are systematicly recieved right before the emmision of post-spikes are increased, and decreased if vice versa. After such training, the neurons begins to actively react to pre-spikes received from neurons with larger connection weights. This leads to the neuron weights learning certain features (Fig. \ref{fig:training}). If these pre-spikes are recieved in a short enough time period, they might cause a new spike. The opposite rule (with $A_{+}$ and $A_{-}$ having the opposite signs) is called the anti-STDP \cite{anti-STDP} rule. It is used to train inhibitory connenctions in section \ref{section:competition-training} of this paper.

@@

\begin{figure}
\centering
\begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{weights_XY_untrained.pdf} 
    \label{fig:training:before_training}
    \caption{}
\end{subfigure}
\begin{subfigure}{0.45\textwidth}  \label{weights_XY}
    \includegraphics[width=\textwidth,keepaspectratio=true]{weights_XY.pdf}
    \caption{} 
\end{subfigure}
\caption{Visualization of $XY$ connection weights of a 900 $Y$ neurons (100 neurons for each of the 9 patches) network --- a 360 $ \times $ 360 matrix. The weights of one neuron correspond to a 12 $\times$ 12 square. The weights of the $Y$ neurons are grouped by patch - the lower left 120 $  \times $ 120 square  (10 $ \times $ 10 neurons) corresponds to the lower left patch, the center square corresponds to the center patch, and so on. Before the trainig process (Fig. (a)) the weights are random. It can be seen that after 1000 training iterations (Fig. (b)) the neurons learn different features, resembling digits elements.}
\label{fig:training}
\end{figure}

@@

For all networks, this calibration process was used for 10,000 examples from a calibration sample especially selected from the test part of the MNIST dataset, as noted above. Note that calibration is nota  part of the network training, since it is only included in the algorithm for interpreting the behavior of the SNS. The votes are used as a measure of confidence of the neuron in each of the digits classes. The example votes are shown on Fig. \ref{fig:votes}.

\begin{figure}
\centering
\begin{subfigure}{\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{votes.pdf}
    \caption{$Y$ neurons votes.}
\end{subfigure}
\begin{subfigure}{\textwidth} 
    \includegraphics[width=\textwidth,keepaspectratio=true]{votes_distr_4.pdf} 
    \caption{Votes distrivution for the 4 most relevant classes (for each neuron). It can be seen that the vote value (average number of spikes per class) drops significantly with relevance decreasing.}
\end{subfigure}
\caption{Visualization of the votes of the $ Y $ network layer with 225 $ Y $ neurons. High values correspond to a large specialization of the neuron in the digits class. It can be seen that almost all neurons have a nonzero vote value, therefore, almost all neurons in the network are active.}
\label{fig:votes}
\end{figure}

@@

Accuracy learning curves (Fig. \ref{fig:learning_curves}) were built for various interpretation algorithms. The accuracy was measured every 250 training iterations. To calibrate the interpretation algorithm at each step a calibration sample of 5000 was used, while to measure the accuracy, a test sample of 1000 was used.

@@

\begin{figure}
\centering
\begin{subfigure}{0.48\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{LCSNN_learning_rate.pdf}
 \caption{LCSNN learning curve.}
 \label{LCSNN_learning_curve}
\end{subfigure} 
\begin{subfigure}{0.48\textwidth} 
    \includegraphics[width=\textwidth,keepaspectratio=true]{CSNN_learning_rate.pdf}
 \caption{CSNN learning curve.}
 \label{CSNN_learning_curve}
\end{subfigure} 
\end{figure}
\begin{figure}\ContinuedFloat
\begin{subfigure}{0.48\textwidth} 
    \includegraphics[width=\textwidth,keepaspectratio=true]{FCSNN_learning_rate.pdf}
 \caption{FCSNN learning curve.}
 \label{FCSNN_learning_curve}
\end{subfigure}
\caption{Comparison of learning curves of different architectures of SNNs with competition: (\ref{LCSNN_learning_curve}) LCSNN, (\ref{CSNN_learning_curve}) CSNN, (\ref{FCSNN_learning_curve}) FCSNN. The results are evaluated with different algorithms for the interpretation of activity. The \textbf{preselections by spikes} algorithm was only checked for LCSNN, because it achieves practically same results as \textbf{patch voting}. Standard deviations are plotted as errors. The networks have 100 channels (for a FCSNN the channel consists of a single neuron), the convolution kernel for LCSNN is 12, and for CSNN it is 8.}
\label{fig:learning_curves}
\end{figure}

\subsection{Comparison of convolution and local connections effectiveness}
Experiments were carried out to measure the accuracy of networks with different architectures. For convolutional and fully connected networks,  inhibitory connections similar to LCSNN were introduced. Note that using a linear classifier as an interpretation algorithm, it was possible to achieve 95\% accuracy for a locally connected network of 1000 channels with kernel size of 12.

@@

@@

\begin{figure}
\centering
\begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_distribution_worst.pdf}
    \caption{Weakest competition}
\end{subfigure}
\begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_distribution_medium_bad.pdf}
    \caption{Weak competition}
\end{subfigure}
\\
\begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_distribution_medium_good.pdf}
    \caption{Medium competition} 
\end{subfigure}
\begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_distribution_best.pdf}
    \caption{Strong competition}
    \label{fig:best_competition}
\end{subfigure}
\caption{Different competition weights distributions}
\label{fig:competition_distributions}
\end{figure}

@@

\begin{figure}
\centering
\begin{subfigure}{0.45\textwidth} 
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_on_XY_worst.pdf}
    \caption{Weakest competition}
    \label{fig:worst_competition_distribution}
\end{subfigure}
\begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_on_XY_medium_bad.pdf}
    \caption{Weak competition}
    \label{fig:medium_bad_competition_distribution}
\end{subfigure}
\\
\begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_on_XY_medium_good.pdf}
    \caption{Medium competition}
    \label{fig:medium_good_competition_distribution}
\end{subfigure}
\begin{subfigure}{0.45\textwidth} 
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_on_XY_best.pdf}
    \caption{Strong compettion}
    \label{fig:best_competition_XY}
\end{subfigure}
\caption{Visualization of competition weights on top of $ XY $ weights for the networks from Fig. (\ref{fig:competition_distributions}). Competition weights are shown only for one center neuron in each receptive field to avoid cluttering the visualization. The saturated blue color corresponds to the competition weights of large absolute value (the arithmetic mean between the weights $ W_{ij} $ and $ W_{ji} $ is used). It can be seen that similar features compete more strongly with each other than different}
\end{figure}

@@

\begin{figure}
\begin{center}
 \includegraphics[,
 width=\textwidth,keepaspectratio=true]{hyperparams.png}

@@

\begin{figure}
\centering 
\begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_distribution_clamp_low.pdf}
    \caption{Inhibitory connections weights distribution, lower clamp} 
\end{subfigure}
\begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{competition_distribution_clamp_high.pdf}
    \caption{Inhibitory connections weights distribution, higher clamp}
\end{subfigure}
\begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio=true]{weights_XY_clamp_low.pdf}
    \caption{Forward connections weights, lower clamp}
\end{subfigure} 
\begin{subfigure}{0.45\textwidth} 
    \includegraphics[width=\textwidth,keepaspectratio=true]{weights_XY_clamp_high.pdf}
    \caption{Forward connections weights, higher clamp}
\end{subfigure} 
\caption{Impact of limiting the values of inhibitory weights on accuracy. The weights were limited to $ -350 $ on the illustration below and $ -100 $ on the illustration above. The accuracy of a network with similar parameters, but without competition limitation is 82.3\%, its weights $ XY $ are shown in Fig. \ref {fig:best_competition_XY}, and the distribution of its inhibitory weights is shown in Fig. \ref{fig:best_competition}. It can be seen that in both cases more blurred forward propagation weights are trained.
}
\label{fig:compe_clamp}
\end{figure}
